---
title: "Network of all codes based on sections borrowed"
author: "Lincoln Mullen"
date: "September 26, 2015"
---

Earlier we made a network graph of all codes using ratio of matches in a code to code comparison. Now lets make a network graph based on sections borrowed.

```{r message=FALSE}
library("readr")
library("igraph")
library("dplyr")
library("textreuse")
library("stringr")
source("R/best-section-matches.R")
source("R/summarize-borrowings.R")
source("R/extract-date.R")
source("R/extract_code_names.R")
```

Read the data.

```{r}
borrowings <- read_csv("out/scores-all-sections-pairs.csv") %>% 
  rename(section_a = a, section_b = b) %>% 
  mutate(code_a = extract_code_names(section_a),
         code_b = extract_code_names(section_b),
         year_a = extract_date(code_a),
         year_b = extract_date(code_b))
borrowings
```

All the codes in our corpus. Which codes aren't included?

```{r}
codes <- c(borrowings$code_a, borrowings$code_b) %>% unique
all_codes <- Sys.glob("legal-codes/*") %>% 
  str_replace("legal-codes/", "") %>% 
  str_replace("\\.txt", "") %>% 
  extract_code_names()
setdiff(all_codes, codes)
```

All the codes that aren't included are plausibly not connected to the corpus.

Calculate section matches and summarize them for a single code.

```{r}
CA1851 <- best_section_matches("CA1851", scores = borrowings, threshold = 0.2)
CA1851
summarize_borrowings(CA1851)
```

Apply that to all the codes.

```{r}
section_pct <- codes %>% 
  lapply(best_section_matches, scores = borrowings, threshold = 0.2) %>% 
  lapply(summarize_borrowings) %>% 
  bind_rows()
section_pct
```

Do some filtering and munge into an edge list.

```{r}
edges <- section_pct %>% 
  filter(percentage_sections >= 0.05,
         !is.na(match_code)) %>% 
  select(original_code, match_code, weight = percentage_sections)
edges
```

Turn that into a graph.

```{r}
g <- graph.data.frame(edges, directed = TRUE)
set.seed(4821)
l <- layout.fruchterman.reingold(g, weights = E(g)$weight)
par(mar = c(0,0,1,0))
plot(g, edge.width = E(g)$weight * 8, layout = l,
     edge.arrow.size = 0.15, vertex.size = 5)
title("Codes of Civil Procedure by percentage of sections")
```
